---
title: "Module 3"
author: "Sankalp Biswal"
date: "2023-11-26"
output: html_document
---

## 1.Importing Libraries
```{r}
library(pacman)
library(tidyverse)
library(janitor)
library(knitr)

```

## 2.Importing dataset

* Selecting variables of interest
```{r}
property_dataset <- read.csv("fy2023-property-assessment-data.csv")

property_subset <- subset(property_dataset,select = c(CITY,LU,LU_DESC,BLDG_TYPE,LAND_SF,GROSS_AREA,BLDG_VALUE,LAND_VALUE,TOTAL_VALUE,GROSS_TAX,YR_BUILT,YR_REMODEL,ROOF_STRUCTURE,BED_RMS,FULL_BTH,KITCHENS,TT_RMS,RES_FLOOR,OVERALL_COND,HEAT_TYPE,AC_TYPE,PROP_VIEW,NUM_PARKING,FIREPLACES,ZIP_CODE))
```

## 3. Data Cleaning

* We'll clean the data using `IQR()` and replace the out of bound values by 'NA' and then impute 'NA' values by **Median** so that we can preserve the original sample size.

```{r}
clean_numeric_columns <- function(x) {
  # Remove outliers based on IQR
  q1 <- quantile(x, 0.25, na.rm = TRUE)
  q3 <- quantile(x, 0.75, na.rm = TRUE)
  iqr_value <- IQR(x, na.rm = TRUE)
  
  lower_bound <- q1 - 1.5 * iqr_value
  upper_bound <- q3 + 1.5 * iqr_value
  
  x_outliers_removed <- ifelse(x < lower_bound | x > upper_bound, NA, x)
  
  # Impute missing values with the median
  x_cleaned <- ifelse(is.na(x_outliers_removed), median(x, na.rm = TRUE), x_outliers_removed)
  
  return(x_cleaned)
}

# Clean all numeric columns in the dataset
property_subset_clean <- property_subset %>%
  mutate(across(where(is.numeric), clean_numeric_columns))

# Remove rows with any NA values after cleaning
property_subset_clean <- na.omit(property_subset_clean)

property_subset_clean <- property_subset_clean %>%
  mutate(CITY = ifelse(CITY == "" | is.na(CITY), "Unknown", CITY))

property_subset_clean <- property_subset_clean %>%
  mutate(OVERALL_COND = ifelse(OVERALL_COND == "" | is.na(OVERALL_COND), "Unknown", OVERALL_COND))

property_subset_clean <- property_subset_clean %>%
  mutate(PROP_VIEW = ifelse(PROP_VIEW == "" | is.na(PROP_VIEW), "Unknown", PROP_VIEW))

property_subset_clean <- property_subset_clean %>%
  mutate(AC_TYPE = ifelse(AC_TYPE == "" | is.na(AC_TYPE), "Unknown", AC_TYPE))
```
------------------------

## 4. Tests

### Overview

* There are 2 parts namely - **PART-1 and PART-2**.
* We'll be conducting one sample t-test on two variables - **LAND_SF** and **GROSS_TAX** for **PART-1** and **PART-2** respectively.
* PART-1 : We'll conduct t-test on 3 samples with varying size for the variable **LAND_SF** and make a decision using t-value for each of the sample.
* PART-2 : We'll conduct t-test on 3 samples with varying size for the variable **GROSS_TAX** and make a decision using p-value for each of the sample. 

------------------------

### **PART-1**

### One sample t-test for LAND_SF i.e land area in sqft using t- statistic

* We'll take 3 samples of different sizes such as 100, 1000, and 10000
* We'll conduct t-test for the 3 samples and analyse if the decision differs.

#### a. Calculation of True Population Mean

* Since the sample that we are testing is not small, we don't need to check for normality

```{r}
# Loading the required variable for test in a separate variable

land_sf <- property_subset_clean$LAND_SF

# Calculating the true value of mean from population

mean_land_sf <- mean(land_sf)

cat("Population mean =",mean_land_sf)
```

#### b. Hypothesis test for Sample 1 (Land sf)

1. Null Hypothesis (H0): The mean land square footage is equal to **2689.257**
2. Alternative Hypothesis (H1): The mean land square footage is not equal to **2689.257**

* Now we'll create a sample and check for the above hypothesis using `t.test()`.

```{r}

set.seed(123)

# Create a random sample of size 100 

sample_size1 <- 100

random_sample1 <- sample(land_sf, size = sample_size1, replace = TRUE)

# Perform one-sample t-test
t_test_result_1 <- t.test(random_sample1, mu = mean_land_sf)

# Display the results
print(t_test_result_1)

# Calculation of t(critical)
confidence_level <- 0.95  
degrees_of_freedom_1 <- 99

# Calculate the critical t-value
t_critical_1 <- qt((1 + confidence_level) / 2, df = degrees_of_freedom_1)

# Print the result

cat("t(critical) for 95% confidence interval and 99 degree of freedom = ",t_critical_1)

```
#### c. Interpretation

* Criterias for rejecting null hypothesis using **t- value**

1. If t > t(critical), Reject Null Hypothesis
2. If t < t(critical), Fail to Reject Null Hypothesis

* t-value: We got t-value = 1.12 and t(critical) = 1.984, t < t(critical), therefore we **fail to reject null hypothesis**.
* Confidence interval: The confidence interval provides a range of values within which we can be 95% confident the true population mean lies. Our 95 percent confidence interval is **2477.236 - 3450.824**. This further strengthens our conclusion of **failing to reject Null hypothesis**

#### d. Decision

* At a 5% significance level or a 95 percent confidence level, we **do not have enough evidence to claim that the sample mean is different from the hypothesized value of 2689.257**. The confidence interval further supports this conclusion, encompassing the hypothesized value.

------------------------

#### a. Hypothesis test for Sample 2 (Land sf)

1. Null Hypothesis (H0): The mean land square footage is equal to 2689.257
2. Alternative Hypothesis (H1): The mean land square footage is not equal to 2689.257

* Now we'll create a sample and check for the above hypothesis using `t.test()`.

```{r}

set.seed(123)

# Create a random sample of size 1000 

sample_size2 <- 1000

random_sample2 <- sample(land_sf, size = sample_size2, replace = TRUE)

# Perform one-sample t-test
t_test_result_2 <- t.test(random_sample2, mu = mean_land_sf)

# Display the results
print(t_test_result_2)

# Calculation of t(critical)
confidence_level <- 0.95  
degrees_of_freedom_2 <- 999  

# Calculate the critical t-value
t_critical_2 <- qt((1 + confidence_level) / 2, df = degrees_of_freedom_2)

# Print the result

cat("t(critical) for 95% confidence interval and 999 degree of freedom = ",t_critical_2)

```
#### b. Interpretation

* Criterias for rejecting null hypothesis using **t- value**

1. If t > t(critical), Reject Null Hypothesis
2. If t < t(critical), Fail to Reject Null Hypothesis

* t-value: We got t-value = 0.145 and t(critical) = 1.984, t < t(critical), therefore we **fail to reject null hypothesis**.
* Confidence interval: The confidence interval provides a range of values within which we can be 95% confident the true population mean lies. Our 95 percent confidence interval is **2562.911 - 2835.853**. This further strengthens our conclusion of **failing to reject Null hypothesis**

#### c. Decision

* At a 5% significance level or a 95 percent confidence level, we **do not have enough evidence to claim that the sample mean is different from the hypothesized value of 2689.257**. The confidence interval further supports this conclusion, encompassing the hypothesized value.

------------------------

#### a. Hypothesis test for Sample 3 (Land sf)

1. Null Hypothesis (H0): The mean land square footage is equal to 2689.257
2. Alternative Hypothesis (H1): The mean land square footage is not equal to 2689.257

* Now we'll create a sample and check for the above hypothesis using `t.test()`.

```{r}

set.seed(123)

# Create a random sample of size 10000 

sample_size3 <- 10000

random_sample3 <- sample(land_sf, size = sample_size3, replace = TRUE)

# Perform one-sample t-test
t_test_result_3 <- t.test(random_sample3, mu = mean_land_sf)

# Display the results
print(t_test_result_3)

# Calculation of t(critical)
confidence_level <- 0.95  
degrees_of_freedom_3 <- 9999  

# Calculate the critical t-value
t_critical_3 <- qt((1 + confidence_level) / 2, df = degrees_of_freedom_3)

# Print the result

cat("t(critical) for 95% confidence interval and 9999 degree of freedom = ",t_critical_3)

```
#### b. Interpretation

* Critereias for rejecting null hypothesis using t- value

1. If t > t(critical), Reject Null Hypothesis
2. If t < t(critical), Fail to Reject Null Hypothesis

* t-value: We got t-value = 1.21 and t(critical) = 1.960, t < t(critical), therefore we **fail to reject null hypothesis**.
* Confidence interval: The confidence interval provides a range of values within which we can be 95% confident the true population mean lies. Our 95   percent confidence interval is **2673.173 - 2758.053**. This further strengthens our conclusion of **failing to reject Null hypothesis**

#### c. Decision

* At a 5% significance level or a 95 percent confidence level, we **do not have enough evidence to claim that the sample mean is different from the hypothesized value of 2689.257**. The confidence interval further supports this conclusion, encompassing the hypothesized value.

------------------------

### Final conclusion of one sample t-test for LAND_SF

* **We don't have enough evidence to reject out null hypothesis** for any of the 3 samples that we took above. **Therefore, sample mean for the 3 samples is not different from our true population mean.**
* We got the following confidence intervals - **2477.236 - 3450.824 for sample 1**,**2562.911 - 2835.853 for sample 2**, and **2673.173 - 2758.053 for sample 3**. We notice that the confidence interval got narrower as we increased the sample size because increasing the sample size reduces the variability of the sample mean and leads to a more accurate estimation of the population parameter, resulting in a **narrower confidence interval**. 

------------------------

### **PART-2**

### One sample t-test for GROSS TAX using p-value

* We'll take 3 samples of different sizes such as 100, 1000, and 100000
* We'll conduct t-test for the 3 samples and analyse if the decision differs.

#### a. Calculation of True Population Mean

* Since the sample that we are testing is not small, we don't need to check for normality

```{r}
# Loading the required variable for test in a separate variable

gross_tax <- property_subset_clean$GROSS_TAX

# Calculating the true value of mean from population

mean_gross_tax <- mean(gross_tax)

cat("Population mean =",mean_gross_tax)
```

#### b. Hypothesis test for Sample 1 (Gross Tax)

1. Null Hypothesis (H0): The mean gross tax is equal to 6187.053
2. Alternative Hypothesis (H1): The mean gross tax is not equal to 6187.053

* Now we'll create a sample and check for the above hypothesis using `t.test()`.

```{r}

set.seed(123)

# Create a random sample of size 100 

sample_size1 <- 100

random_sample1 <- sample(gross_tax, size = sample_size1, replace = TRUE)

# Perform one-sample t-test
t_test_result_1 <- t.test(random_sample1, mu = mean_gross_tax)

# Display the results
print(t_test_result_1)

```
#### b. Interpretation

* Critereias for rejecting null hypothesi using p-value

1. If p-value <= 0.05 (Standard significance value), Reject Null Hypothesis
2. If p-value > 0.05 (Standard significance value), Fail to reject Null Hypothesis

* p-value: We got p-value = 0.3854 i.e p > 0.05, therefore we **fail to reject null hypothesis**
* Confidence interval: The confidence interval provides a range of values within which we can be 95% confident the true population mean lies. Our 95   percent confidence interval is **5078.313 - 6618.906**. This further strengthens our conclusion of **failing to reject Null hypothesis**

#### c. Decision

* At a 5% significance level or a 95 percent confidence level, we **do not have enough evidence to claim that the sample mean is different from the hypothesized value of 6187.053**. The confidence interval further supports this conclusion, encompassing the hypothesized value.

------------------------

#### a. Hypothesis test for Sample 2 (Gross Tax)

1. Null Hypothesis (H0): The mean land square footage is equal to 6187.053
2. Alternative Hypothesis (H1): The mean land square footage is not equal to 6187.053

* Now we'll create a sample and check for the above hypothesis using `t.test()`.

```{r}

set.seed(123)

# Create a random sample of size 1000 

sample_size2 <- 1000

random_sample2 <- sample(gross_tax, size = sample_size2, replace = TRUE)

# Perform one-sample t-test
t_test_result_2 <- t.test(random_sample2, mu = mean_gross_tax)

# Display the results
print(t_test_result_2)

```
#### b. Interpretation

* Criterias for rejecting null hypothesis using p-value

1. If p-value <= 0.05 (Standard significance value), Reject Null Hypothesis
2. If p-value > 0.05 (Standard significance value), Fail to reject Null Hypothesis

* p-value: We got p-value = 0.9202 i.e p > 0.05, therefore we **fail to reject null hypothesis**
* Confidence interval: The confidence interval provides a range of values within which we can be 95% confident the true population mean lies. Our 95   percent confidence interval is **5925.052 - 6423.594**. This further strengthens our conclusion of **failing to reject Null hypothesis**

#### c. Decision

* At a 5% significance level or a 95 percent confidence level, we **do not have enough evidence to claim that the sample mean is different from the hypothesized value of 6187.053**. The confidence interval further supports this conclusion, encompassing the hypothesized value.

------------------------

#### a. Hypothesis test for Sample 3 (Gross Tax)

1. Null Hypothesis (H0): The mean land square footage is equal to 2689.257
2. Alternative Hypothesis (H1): The mean land square footage is not equal to 2689.257

* Now we'll create a sample and check for the above hypothesis using `t.test()`.

```{r}

set.seed(123)

# Create a random sample of size 100000 

sample_size3 <- 100000

random_sample3 <- sample(gross_tax, size = sample_size3, replace = TRUE)

# Perform one-sample t-test
t_test_result_3 <- t.test(random_sample3, mu = mean_gross_tax)

# Display the results
print(t_test_result_3)

```
#### b. Interpretation

* Criterias for rejecting null hypothesis using p-value

1. If p-value <= 0.05 (Standard significance value), Reject Null Hypothesis
2. If p-value > 0.05 (Standard significance value), Fail to reject Null Hypothesis

* p-value: We got p-value = 0.9744 i.e p > 0.05, therefore we **fail to reject null hypothesis**
* Confidence interval: The confidence interval provides a range of values within which we can be 95% confident the true population mean lies. Our 95   percent confidence interval is **6162.455 - 6212.482**. This further strengthens our conclusion of **failing to reject Null hypothesis**

#### c. Decision

* At a 5% significance level or a 95 percent confidence level, we **do not have enough evidence to claim that the sample mean is different from the hypothesized value of 6187.053**. The confidence interval further supports this conclusion, encompassing the hypothesized value.

------------------------

### Final conclusion of one sample t-test for GROSS_TAX

* **We don't have enough evidence to reject out null hypothesis** for any of the 3 samples that we took above. **Therefore, the sample mean for the 3 samples are not different from the true population mean**
* We got the following confidence intervals - **5078.313 - 6618.906 for sample 1**,**5925.052 - 6423.594 for sample 2**, and **6162.455 - 6212.482 for sample 3**. We notice that the confidence interval got narrower as we increased the sample size because increasing the sample size reduces the variability of the sample mean and leads to a more accurate estimation of the population parameter, resulting in a **narrower confidence interval**. 

-------------------------














